---
title: "Plots locus"
author: "Athena Tamayo Luisce"
date: "3/6/2021"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
setwd("/home/athena/Escritorio/CCG/4toSemestre/GenoHumana/Proyecto")
data1 <- read.table("./ANNOVAR_hg38_filtered/Monogenic3.variant_function", sep = "\t", header = FALSE)
data2 <- read.table("./ANNOVAR_hg38_invalid_filtered/Monogenic3_1.variant_function", sep = "\t", header = FALSE)
exons1 <- read.table("./ANNOVAR_hg38_filtered/Monogenic3.exonic_variant_function", sep = "\t", header = FALSE)
exons2 <- read.table("./ANNOVAR_hg38_invalid_filtered/Monogenic3_1.exonic_variant_function", sep = "\t", header = FALSE)
```

Merge data1/data2 and exon1/exon2
```{r}
data <- left_join(data1, data2)
exons <- left_join(exons1, exons2)
```

```{r}
dim(data)
dim(exons)
```

Save the number of loci in a variable 
```{r}
nrowData <- nrow(data)
nrowExons <- nrow(exons)
```

```{r}
head(exons)
```
```{r}
head(exons)
```

Change column names 
```{r}
colnames(data) <- c("location", "gene", "description")
colnames(exons) <- c("line", "mutation", "gene", "description")
```

Create dataframe of frequencies 
```{r}
freq <- as.data.frame(table(data$location))
# add the percentages 
freq$Perc <- round((freq$Freq)/sum(freq$Freq)*100, 2)
```

Plot the frequency of mutations associated with monogenic diseases located in different genomic regions.
```{r}
ggplot(freq, aes(x = Var1, y = log(Freq), fill = Var1)) +
  geom_bar(stat = "identity") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.05),
        legend.title =  element_text(face = "bold")) +
  labs(title = "Frequency of loci associated with monogenic diseases located 
                            in different genomic regions", x = NULL, y = "log(frequency)") +
  scale_fill_discrete(name = "Genomic location") +
  geom_text(aes(label = paste(Perc, "%", sep = "")), vjust = -0.2) 
```
Label info: https://annovar.openbioinformatics.org/en/latest/user-guide/gene/

Plot the frequency of the types of the type of mutations associated with monogenic diseases in exons 
```{r}
mut <- as.data.frame(table(exons$mutation))
mut$Perc <- round((mut$Freq)/sum(mut$Freq)*100, 2)
```

```{r}
ggplot(mut, aes(x = Var1, y = log(Freq), fill = Var1)) +
  geom_bar(stat = "identity") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.05),
        legend.title =  element_text(face = "bold")) +
  labs(title = "Frequency of the type of mutations that gave rise to the exonic loci\n                       associated with monogenic diseases",
       x = NULL, y = "log(frequency)") +
  scale_fill_discrete(name = "Type of mutations") +
  geom_text(aes(label = paste(Perc, "%", sep = "")), vjust = -0.2) +
  ylim(0, 10)
```

Split the description of the variants 
```{r}
elements <- strsplit(data$description, split = "[,]")
```

Plot the number of diseases associated with a number of variants
```{r}
disease <- vector()
for (i in 1:nrowData){
  disease <- c(disease, elements[[i]][4])
}
```

```{r}
disease <- as.data.frame(table(disease))
```

```{r}
ggplot(disease, aes(x = Freq)) + 
  geom_histogram(binwidth = 1, color = "black", fill = "darkseagreen1") +
  labs(title = "Number of genes with a certain number of loci associated \nwith monogenic diseases",
       x = "Number of loci", y = "Number of genes") +
  theme(plot.title = element_text(face = "bold")) +
  scale_x_continuous(breaks = seq(1, 32, 2), limits = c(0, 30))
```
```{r}
summary(disease$Freq)
```

```{r}
table(disease$Freq)
```
```{r}
# Obtain the mim number of the disease that is associated with 166 variants 
print(which(disease$Freq == 75))
disease$disease[311]
```
Plot the distribution of the mutations on each chromosomes
```{r}
desc <- strsplit(data$description, split = "[ ]")
chrs <- vector()
for (i in 1:nrowData){
   chrs <- c(chrs, desc[[i]][1])
}

chrs <- as.data.frame(table(chrs))
order <- order(c(1,10,11,12,13,14,15,16,17,18,19,2,20,21,22,3,4,5,6,7,8,9,23,24))
```

```{r}
chrs <- data.frame(chrs = chrs$chrs[order], Freq = chrs$Freq[order])
chrs$chrs <- factor(chrs$chrs, unique(chrs$chrs))
```

```{r}
palette <- grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
colors <- sample(palette, 24, replace = FALSE)

ggplot(chrs, aes(x = "", y = Freq, fill = chrs)) +
  geom_bar(stat = "identity", color='black') +
  coord_polar("y", start = 0) +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        panel.grid  = element_blank(),
        plot.title = element_text(face = "bold"),
        legend.title =  element_text(face = "bold"))+
  scale_fill_discrete(name = "Chromosomes", type = colors) +
  labs(title = "Distribution of the mutations on each chromosome", x = "", y = "") 
```

```{r}
chrs$Perc <- round(chrs$Freq/sum(chrs$Freq)*100, 2)
```

```{r}
ggplot(chrs, aes(x = chrs, y = Freq, fill = chrs)) +
  geom_bar(stat = "identity") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x=element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.05),
        legend.title =  element_text(face = "bold")) +
  labs(title = "Frequency of loci associated with monogenic diseases in each\nchromosome",
       x = NULL, y = "Frequency") +
  scale_fill_discrete(name = "  Type of\nmutations") +
  geom_text(aes(label = paste(Perc, "%", sep = "")), vjust = -0.2) +
  ylim(c(0, 2000))
```